---
title: "Principles of Statistical Analysis Report"
author: 
- Robin Boudry, Lin Tang, Mengistu Meron, Stijn...
date: "2020-01-03"
output: 
  pdf_document:
    fig_caption: yes
    highlight: pygments
    keep_tex: yes
    number_sections: yes
    toc: no
header-includes: 
  - \usepackage{float}
  - \usepackage{longtable}
---
\centering

\raggedright

\clearpage

\tableofcontents


```{r, include=FALSE}
#INFO
#5 pages MAX
# no intro needed, just METHODS, RESULTS, CONCLUSION
# mention if assumptions are met for each test

```


# Methods

## Data Cleaning & Transformation
Data of 40 subjects was provided, containing information of relative abundances of 8 bacterial species (to be divided in 2 geni) present in their armpit, their BMI, age and gender. 1 observation was removed due to missing data. The categorical variable 'Gender' needed some cleaning to remove whitespaces around some codings of gendertype. For readability and easier coding the names 'Corynebacterium1' etc.. was shortened to 'Cb1' as was 'Staphylococcus1' etc.. shortened to 'St1'
A string variable 'top3' was created, refering to the top 3 classes per individual.
This gives us a 26 categories for 39 individuals...We also create top1 and top2, further aggregating the dominant bacteria and lessening the possible combinations.


### Descriptive analysis
When we view the data of age and CBorST, and sort on CBorST, we see no one above 30 has CB as dominant group. First indication of change in composition. When plotting the Geni with age, we see Corynebacterium having a cluster of low values in the younger age group, and Staphylococcus having a cluster of high values in the younger age group. 

### Categorical variable
Moreover, the continous variable Age is discreted to Agecat as a categorical variable The Agecat is a 3 level ordered factor variable made up by Junior(<=26), Adult(27<=age<=50), Senior(>=51), which is prepared for analysis the relationship of agecat and genus composition.

## Categorical variable analysis
### Interaction among factors
We should confirm that there is no interaction among those factors(Agecat, Gender, BMI) before we do the ANOVA for each of them. We used the two-way ANOVA and interaction graphs to check if the intercations are statistically significant. If the the interactions between Agecat and other two factors are statistically significant, the Agecat variable cannot be analyzed. In this situation, we have to take consider of the interaction of both factors rather than the individual factor. If there are no interaction between each two of them, then we can do the one-way ANOVA to Agecat.

### One-way ANOVA
The sample sizes of each Agecat are unequal, the normally one-way ANOVA and POST HOC test, such as Fisher's LSD and Tukey's HSD, are not availiable for this situation. So we choosed the Kruskal–Wallis test with the null hypothesis that the medians of all groups are equal, and the alternative hypothesis that at least one population median of one group is different from the population median of at least one other group.


```{r setup, echo=FALSE, message=FALSE, include=FALSE}
#'Setting up data & small cleaning & transformation'

#Pre Explorative: glancing
##################
data <- read.csv(paste0(getwd(),"/data/armpit.txt"), sep = ' ')
data <- data[complete.cases(data),]
data$Gender <- factor(trimws(data[["Gender"]], which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))
##################

#Rename the vars for readability
##################
names(data) <- c("Cb1", "Cb2", "Cb3", "Cb4", 
                 "St1", "St2", "St3", "St4",  
                 "Age", "BMI", "Gender" )
##################

#Create top 3 classes per observations
##################
#x <- toString(names(sort(data[1,1:8],decreasing = TRUE)[1:3]))
#top3
data$top3 <- 0
for(i in 1:nrow(data)){
  data$top3[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:3]))
}
#table(data$top3)

#We see 26 combinations for 39 observations.... With this data any difference
#test will seem significant.
#Let's build top 2

#top2
data$top2 <- 0
for(i in 1:nrow(data)){
  data$top2[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:2]))
}
#table(data$top2)
#We still see 13 categories here... Still a lot.
#Lets build top 1

#top1
data$top1 <- 0
for(i in 1:nrow(data)){
  data$top1[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1]))
}
#table(data$top1)
#Here we have only 4 categories. However the question was composition.
#Let's aggregate Cv and ST, and then build top 2 & top 3

#aggregation
data$CB <- 0
data$ST <- 0
for(i in 1:nrow(data)){
  data$CB[i] <- sum(data[i,1:4])
  data$ST[i] <- sum(data[i,5:8])
}
par(mfrow=c(1,2))
plot(data$Age, data$CB, main = 'CB & Age')
plot(data$Age, data$ST, main = 'ST & Age')
par(mfrow=c(1,1))
#We get a hint of increase in CB, and decrease in ST, but needs testing.

#Let's build CB_ST:
#Which is dominant?
data$CBorST <- 0
for(i in 1:nrow(data)){
  data$CBorST[i] <- toString(names(sort(data[i,15:16],decreasing = TRUE)[1:2]))
}
#table(data$CBorST)

#When we view the dataset, and sort on CBorST, we see no one above 30 has CB
#as dominant group. First indication of change in composition
#shown with dplyr
library(dplyr)
#data%>%
#  select(Age, CBorST)%>%
#  filter(CBorST == "CB, ST")

#Age discrete to Agecat as 3 ordered factors
data <- transform(data, Agecat = Age)
data <- within(data,{
                    Agecat[Age <= 25] <- "Junior"
                    Agecat[Age >= 26] <- "Adult"
                    Agecat[Age >= 51] <- "Senior"
                    })
data$Agecat <- factor(data$Agecat, ordered = TRUE, levels = c('Junior','Adult', 'Senior'))

##################


```


## Choice of non parametric tests
We are confronted with asample size and since no information is given about the underlying distribution, we have to rely on non-parametric methods to make valid inferences.
For correlation testing, we consulted Hauke & Kowoski(2011)'s comparison of correlation tests, and concluded that Spearman rank was suitable for our problem.

To better understand the Spearman Rank correlation significance test, we simulated
data with the same sample size (n1=n2=13) which resembles what we observed and with higher sample sizes.. We noted that smaller sample sizes take higher correlation estimates for a significant p-value.
Here we will sample from the uniform distribution and multiply by 100 (reflects our data)
For Type1 testing, we sample from the exact distribution
For Power testing, we limit the var1 range between .40 & .55, and create var2 by substracting var1 from 100 so they sum to 100 (as in our aggregated CB & ST).

We also checked the Type1 error in a simulation, and saw it being controlled for Spearman Rank cor test, similar to more established tests. 

For group comparisons the anova was out of order, but the Wilcoxon Rank test as seen in the course was suitable.


# Results

## Agecat analysis
### Interaction analysis

The interaction graphs both show the across of the lines, which means there are interactions between Agecat and Gender/BMI. From the ANOVA tables, we found the p-value of Agecat:Gender and Agecat:BMI are 0.3689 and 0.2056 perspectively. As the interactions are not statistically significant, we are allowed to use one-way ANOVA to analysis the Agecat varible individully.
```{r echo=FALSE, message=FALSE}

#Interaction graphs
attach(data)
par(mfrow=c(1,2))
interaction.plot(Agecat, Gender, ST, 
                 xlab = "Agecat", 
                 ylab = "Staphylococcus", 
                 trace.label = "Gender",
                 main = "Interaction of Agecat and Gender",
                 ylim = c(0,100),  
                 pch = c(19,17), 
                 fixed = TRUE)
interaction.plot(Agecat, BMI, ST, 
                 xlab = "Agecat", 
                 ylab = "Staphylococcus", 
                 trace.label = "BMI",
                 main = "Interaction of Agecat and BMI",
                 ylim = c(0,100),  
                 pch = c(19,17), 
                 fixed = TRUE)

#Two-way ANOVA
#Agecat vs Gender
fit_1 <- aov(ST~Agecat*Gender)
summary(fit_1)
#Agecat vs BMI
fit_2 <- aov(ST~Agecat*BMI)
summary(fit_2)

```

### Kruskal-Wallis test of Agecat

The boxplot shows that there are some difference between 3 agecats. From the result of the Kruskal-Wallis rank sum test, we found the p-value is 0.09773, which is greater than alpha(0.05), so we cannot reject the null hypothesis with this test.
```{r echo=FALSE, message=FALSE}

#Kruskal Wallis test
boxplot(ST~Agecat,
        ylab = "Staphylococcus",
        main = "Staphylococcus by Agecat")
kruskal.test(ST~Agecat)
```
## Age as continuous variable

To verify if genus composition changes with age we use smooth line plot for each species in the two genus groups. However it is noticed that availability of relative abundances of all four species are not equal. For instance,  from Corynebacteria, Corynebacterium.2 is most dominant species, followed by Corynebacterium.1 (see appendix... for detail plots). In the Corynebacterium.1, the relative abundancy keeps constant value of 0% at young ages and suddenly increases  at around 30 year. It reaches maximum (approximately 41%)in age between 40 and 60 years, where after 60 year it gradually drops to 0%. 

In Staphylococcus genus the dominant species is staphylococcus.1 (appendixx..). In opposite to Corynebacteria at young age it attains higher relative abundance (greater than 75%), which gradually decreases till age of 40 year. The relative abundance remains constant at its minimum (below 50 ) between age of 40 and 60 years, and increases for ages older than 60 year.

```{r include=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)
# install.packages("gridExtra")
library(gridExtra)
#setwd("/Users/meron/Downloads/2019_2020/prinstat/project")
#setwd("/Users/meron/Downloads/2019_2020/prinstat/project")
data <- read.table("armpit.txt", header=TRUE, sep='')
summary(is.na(data)) # 1 found in age which is a similar observation for gender too

#cleaning missing values
data <-na.omit(data)
summary(is.na(data)) # no missing values

#creat variables for Corynebacterium species
coryn1<-data$Corynebacterium.1
coryn2<-data$Corynebacterium.2
coryn3<-data$Corynebacterium.3
coryn4<-data$Corynebacterium.4

#total Corynebacterium by adding the four species
data$corynTot<-(coryn1+coryn2+coryn3+coryn4)
Corynebacterium<-data$corynTot

#creat variables for Staphylococcus species
staph1<-data$Staphylococcus.1
staph2<-data$Staphylococcus.2
staph3<-data$Staphylococcus.3
staph4<-data$Staphylococcus.4
##total Staphylococcus by adding the four species
data$staphTot<-staph1+staph2+staph3+staph4
Staphylococcus<-data$staphTot

age<-data$Age
gender<-data$Gender
BMI<-data$BMI
##################
```

 
```{r echo = FALSE}
###agregated species smoth line plots
ggplot(data = data, aes(x = age, y = Staphylococcus)) +
  geom_point() +
  geom_smooth()

ggplot(data = data, aes(x = age, y = Corynebacterium)) + 
  geom_point() +
  geom_smooth()
```

For ease of interpretation and  comparison between  two Genus groups, we aggregate each of the four species in to one respective  group,  Corynebacterium and Staphylococcus. The result shows that the distribution of aggregated genus groups follows the dominant species distribution. As a result for Corynebacterium the relative abundancy gets minimum at the early age of around 20, and suddenly increases at around 24 year. It gets maximum approximately at age 43 and 60 years.  It drops gradually for ages older than 60 year. 

In case of Staphylococcus the relative abundancy is high at early age of about 20 and gradually deacreses with increasing age.  Lower relative abundacy for this genus is attined when age is between 40 and 60.  For older ages than 60 its availability decreases so as its amount.

It is clearly seen that the change in relative abundancy with age is not linear. consquently the correlations between the aggregated relative abundancy and age is calculated using the spearman correlation test in R.  correlation coefficient obtained for Corynebacterium and Staphylococcus is 0.40 and -0.41.
```{r include=FALSE}

#correlation test
cor.test(Corynebacterium, age, method = "spearman")
cor.test(Staphylococcus, age, method = "spearman")


```


```{r echo = FALSE}
#### plot for all 8 species 

pcor1<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Corynebacterium.1))

pcor2<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Corynebacterium.2))

pcor3<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Corynebacterium.3))

pcor4<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Corynebacterium.4))

pstaph4<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Staphylococcus.4))

pstaph3<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Staphylococcus.3))

pstaph2<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Staphylococcus.2))

pstaph1<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Staphylococcus.1))
```


# Conclusion

## Genus composition relate to age

### Genus composition among agecat
Since the p-value is greater than 0.05, we cannot to reject the null hypothesis, which means the genus composition do not shows significant difference among Agecat. 

# References
Hauke, J., & Kossowski, T. (2011). Comparison of values of Pearson's and Spearman's correlation coefficients on the same sets of data. Quaestiones geographicae, 30(2), 87-93.

# Appendix
plots for each species versus age continuous
```{r echo = FALSE}
grid.arrange(pcor1, pcor2, pcor3, pcor4, nrow = 2)
grid.arrange(pstaph1, pstaph2, pstaph3, pstaph4, nrow = 2)
```
A look at difference of Spearman Rank correlation estimate significanes for different sample sizes
```{r spearmanbehav, echo=FALSE, message=FALSE}
p <- est <- px <- estx <- c()
for (i in 1:10000){
  Y1 <- rnorm(13)
  Y2 <- rnorm(13)
  Y1x <- rnorm(60)
  Y2x <- rnorm(60)
  p[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  est[i] <- cor.test(Y1,Y2, method = "spearman")$estimate
  px[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
  estx[i] <- cor.test(Y1x,Y2x, method = "spearman")$estimate
}
x <- data.frame(est, p)
xx <- data.frame(estx, px)
par(mfrow=c(1,2))

hist(x[x$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(x[x$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=13)'))

hist(xx[xx$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(xx[xx$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=60)'))

par(mfrow=c(1,1))
```

Type1 simulation for Spearman rank.
```{r spearmanType1, echo=FALSE, message=FALSE}

#H0 & Power testing for our groups
#assuming an even n in groups of 13 (39/3)
#sampling from runif and *100 to reflect out data
#Working with the aggregated level of CB & ST
##################
library(coin)
results_H0_A_runif <- data.frame(size=c(rep(26,6)),dist=rep('runif',6),
                               test=rep(c('perm_t','WMW','Spearman R'),2),
                               Test=c(rep('Type1',3),rep('Power',3)),
                               Value= 0)
pt_p <- wmw_p <- spr_p <- c()
p.t <- p.wmw <- p.spr <- c()
pt_pA <- wmw_pA <- spr_pA <- c()
p.tA <- p.wmwA <- p.sprA <- c()
n <- 13
for (i in 1:1000){
  #For H0= mean from same distrib
  Y1 <- runif(n)*100
  Y2 <- runif(n)*100
  #Groups
  X <- factor (c( rep("CB",n), rep("ST",n )))
  Y <- c(Y1 , Y2)
  #Power, now we need an actual difference
  Y1x <- runif(n, min = .40, max = .55)*100
  Y2x <- 100-Y1x
  #Groups
  Xx <- factor (c( rep("CB",n), rep("ST",n )))
  Yx <- c(Y1x , Y2x)
  # permutation H0
  p.t[i] <- pvalue(oneway_test(Y ~ X,
                                distribution = approximate (nresample =9999)))
  p.wmw[i] <- wilcox.test(Y1 ,Y2 ,
                          exact = TRUE )$p.value
  p.spr[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  # permutation HA
  p.tA[i] <- pvalue(oneway_test(Yx ~ Xx,
                               distribution = approximate (nresample =9999)))
  p.wmwA[i] <- wilcox.test(Y1x ,Y2x ,
                          exact = TRUE )$p.value
  p.sprA[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
}
# Monte - Carlo approximation of the power
# for alpha = .05


results_H0_A_runif$Value[1] <- mean(p.t < .05)
results_H0_A_runif$Value[2] <- mean(p.wmw < .05)
results_H0_A_runif$Value[3] <- mean(p.spr < .05)  
results_H0_A_runif$Value[4] <- mean(p.tA < .05)
results_H0_A_runif$Value[5] <- mean(p.wmwA < .05)
results_H0_A_runif$Value[6] <- mean(p.sprA < .05) 

library(ggplot2)
ggplot(results_H0_A_runif, aes(factor(dist), Value, fill = test)) + 
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Greys")+
  geom_hline(yintercept=0.05, col = 'red')+
  ggtitle("Power, Type1, for 10k reps")+
  facet_grid(. ~ Test)+
  theme_classic()
```

