---
title: "Principles of Statistical Analysis Project Report"
author: "Robin Boudry, Lin Tang, Mengistu Meron, Stijn Vandenbulcke"
date: '2020-01-03'
output:
  pdf_document:
    fig_caption: yes
    highlight: pygments
    keep_tex: yes
    number_sections: yes
    toc: no
  html_document:
    df_print: paged
    toc: no
header-includes:
- \usepackage{float}
- \usepackage{longtable}
---
\centering

\raggedright

\clearpage

\tableofcontents

\clearpage


```{r, include=FALSE}
#INFO
#5 pages MAX
# no intro needed, just METHODS, RESULTS, CONCLUSION
# mention if assumptions are met for each test
```


# Methods

## Data Cleaning & Transformation
Data of 40 subjects was provided, containing information of relative abundances of 8 bacterial species (to be divided in 2 geni) present in their armpit, their BMI, age and gender. 1 observation was removed due to missing data. The categorical variable 'Gender' needed some cleaning to remove whitespaces around some codings of gendertype. For readability and easier coding the names 'Corynebacterium1' etc.. was shortened to 'Cb1' as was 'Staphylococcus1' etc.. shortened to 'St1'
A string variable 'top3' was created, refering to the top 3 classes per individual.
This gives us a 26 categories for 39 individuals...We also create top1 and top2, further aggregating the dominant bacteria and lessening the possible combinations.


### Descriptive analysis
When making boxplots of the species (appendix) we see the relative abundance of Staphylococcus 1 being obviously dominant with a mean of 71.21 and a median of 93.67. This skewnedd is somewhat mediated by using a log transformation. There are 24 females and 15 males in the sample.
When we view the data of age and CBorST, and sort on CBorST, we see no one above 30 has CB as dominant group and takes this as a first indication of change in bacteria composition with age. When plotting the Geni with age, we see Corynebacterium having a cluster of low values in the younger age group, and Staphylococcus having a cluster of high values in the younger age group. 

### Categorical variable
Moreover, the continous variable Age is discreted to Agecat as a categorical variable The Agecat is a 3 level ordered factor variable made up by Junior(<=26), Adult(27<=age<=50), Senior(>=51), which is prepared for analysis the relationship of agecat and genus composition.

## Categorical variable analysis
### Interaction among factors
We should confirm that there is no interaction among those factors(Agecat, Gender, BMI) before we do the ANOVA for each of them. We used the two-way ANOVA and interaction graphs to check if the intercations are statistically significant. If the the interactions between Agecat and other two factors are statistically significant, the Agecat variable cannot be analyzed. In this situation, we have to take consider of the interaction of both factors rather than the individual factor. If there are no interaction between each two of them, then we can do the one-way ANOVA to Agecat.

### One-way ANOVA
The sample sizes of each Agecat are unequal, the normally one-way ANOVA and POST HOC test, such as Fisher's LSD and Tukey's HSD, are not availiable for this situation. So we choosed the Kruskal–Wallis test with the null hypothesis that the medians of all groups are equal, and the alternative hypothesis that at least one population median of one group is different from the population median of at least one other group.


```{r setup, echo=FALSE, message=FALSE, include=FALSE}
#'Setting up data & small cleaning & transformation'
#Pre Explorative: glancing
##################
data <- read.csv(paste0(getwd(),"/data/armpit.txt"), sep = ' ')
data <- data[complete.cases(data),]
data$Gender <- factor(trimws(data[["Gender"]], which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))
##################
#Rename the vars for readability
##################
names(data) <- c("Cb1", "Cb2", "Cb3", "Cb4", 
                 "St1", "St2", "St3", "St4",  
                 "Age", "BMI", "Gender" )
##################
#Create top 3 classes per observations
##################
#x <- toString(names(sort(data[1,1:8],decreasing = TRUE)[1:3]))
#top3
data$top3 <- 0
for(i in 1:nrow(data)){
  data$top3[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:3]))
}
#table(data$top3)
#We see 26 combinations for 39 observations.... With this data any difference
#test will seem significant.
#Let's build top 2
#top2
data$top2 <- 0
for(i in 1:nrow(data)){
  data$top2[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:2]))
}
#table(data$top2)
#We still see 13 categories here... Still a lot.
#Lets build top 1
#top1
data$top1 <- 0
for(i in 1:nrow(data)){
  data$top1[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1]))
}
#table(data$top1)
#Here we have only 4 categories. However the question was composition.
#Let's aggregate Cv and ST, and then build top 2 & top 3
#aggregation
data$CB <- 0
data$ST <- 0
for(i in 1:nrow(data)){
  data$CB[i] <- sum(data[i,1:4])
  data$ST[i] <- sum(data[i,5:8])
}
par(mfrow=c(1,2))
plot(data$Age, data$CB, main = 'CB & Age')
plot(data$Age, data$ST, main = 'ST & Age')
par(mfrow=c(1,1))
#We get a hint of increase in CB, and decrease in ST, but needs testing.
#Let's build CB_ST:
#Which is dominant?
data$CBorST <- 0
for(i in 1:nrow(data)){
  data$CBorST[i] <- toString(names(sort(data[i,15:16],decreasing = TRUE)[1:2]))
}
#table(data$CBorST)
#When we view the dataset, and sort on CBorST, we see no one above 30 has CB
#as dominant group. First indication of change in composition
#shown with dplyr
library(dplyr)
#data%>%
#  select(Age, CBorST)%>%
#  filter(CBorST == "CB, ST")
#Age discrete to Agecat as 3 ordered factors
data <- transform(data, Agecat = Age)
data <- within(data,{
                    Agecat[Age <= 25] <- "Junior"
                    Agecat[Age >= 26] <- "Adult"
                    Agecat[Age >= 51] <- "Senior"
                    })
data$Agecat <- factor(data$Agecat, ordered = TRUE, levels = c('Junior','Adult', 'Senior'))
##################
```


## Relation of 8 relative abundances amongst each other
### Choice of test
We are confronted with a small sample size and since no information is given about the underlying distribution, we have to rely on non-parametric methods to make valid inferences.
For correlation testing, we consulted Hauke & Kowoski(2011)'s comparison of correlation tests, and concluded that Spearman rank was suitable for our problem.

To better understand the Spearman Rank correlation significance test, we simulated
data with the same sample size (n1=n2=13) which resembles what we observed and with higher sample sizes.. We noted that smaller sample sizes take higher correlation estimates for a significant p-value.
We also checked the Type1 error in a simulation, and saw it being controlled at the specified 0.05 level for our Spearman Rank cor test, similar to more established tests. 

### Hypothesis and testing procedure

The null hypothesis can be stated as follows: Spearman's p = 0

This testing can be done with the help of a permutation test. For each combination of species a permutation distribution is made, by permuting the relative abundances and calculating the Spearman correlation. Since the many possible permutations, we rely on the package "Psych", which calculates a p-value via the asymptotic t approximation with the formula:
"t=(r*sqrt(n-2))/sqrt(1-r^2)" and "se=sqrt((1-r²)/(n-2))" 

### Multiple testing adjustment
To adjust for the multiple testing problem, we use the Benjamini-Yekutieli (BY) method, this is a less conservative correction than Bonferroni or Holms-Bonferroni (HB), which gives us more statistical power. However in our case, the amount of significant pairs did not change whether BY or HB was used. 

# Results

## Age categorical (agecat) analysis
### Interaction analysis

The interaction graphs (appendix: interaction of agecat & BMI/Gender) both show the across of the lines, which means there are interactions between Agecat and Gender/BMI. From the ANOVA tables, we found the p-value of Agecat:Gender and Agecat:BMI are 0.37 and 0.21 perspectively. As the interactions are not statistically significant, we are allowed to use one-way ANOVA to analysis the Agecat variable individully.
```{r, include=FALSE}
#Agecat vs Gender
attach(data)
fit_1 <- aov(ST~Agecat*Gender)
summary(fit_1)
#Agecat vs BMI
fit_2 <- aov(ST~Agecat*BMI)
summary(fit_2)
```

Checking the resutls of "aov(ST~Agecat\*Gender)"  we find a p-value of 0.0433 and for "aov(ST~Agecat\*BMI)" a p-value of 0.0384. Hence we can conclude the effect of the categorical age classes stays significant on the presence of Staphylococcus, even when adding Gender and BMI with their respective interactions.

### Kruskal-Wallis test of Agecat

The boxplot (appendix: boxplot of categorical age) shows that there are some difference between 3 agecats. From the result of the Kruskal-Wallis rank sum test, we found the p-value is 0.09773, which is greater than alpha(0.05), so we cannot reject the null hypothesis with this test.
```{r include=FALSE, message=FALSE}
#Kruskal Wallis test

kruskal.test(ST~Agecat)
```
## Age as continuous variable

To verify if genus composition changes with age we use smooth line plot for each species in the two genus groups. However it is noticed that availability of relative abundances of all four species are not equal. For instance,  from Corynebacteria, Corynebacterium.2 is most dominant species, followed by Corynebacterium.1 (see appendix... for detail plots). In the Corynebacterium.1, the relative abundancy keeps constant value of 0% at young ages and suddenly increases  at around 30 year. It reaches maximum (approximately 41%)in age between 40 and 60 years, where after 60 year it gradually drops to 0%. 

In Staphylococcus genus the dominant species is staphylococcus.1 (appendix..). In opposite to Corynebacteria at young age it attains higher relative abundance (greater than 75%), which gradually decreases till age of 40 year. The relative abundance remains constant at its minimum (below 50 ) between age of 40 and 60 years, and increases for ages older than 60 year.

```{r echo = FALSE, message=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
data$age <- data$Age
#### plot for all 8 species 
pcor1<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Cb1))+
  ylab("CB1 Relative Abundancy (%)") + 
  xlab("Age (year)")
pcor2<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Cb2)) +
  ylab("CB2 Relative Abundancy (%)") + 
  xlab("Age (year)")
pcor3<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Cb3)) +
  ylab("CB3 Relative Abundancy (%)") + 
  xlab("Age (year)")
pcor4<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = Cb4)) +
  ylab("CB4 Relative Abundancy (%)") + 
  xlab("Age (year)")
  
pstaph4<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = St4)) +
  ylab("SB4 Relative Abundancy (%)") + 
  xlab("Age (year)")
pstaph3<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = St3)) +
  ylab("SB3 Relative Abundancy (%)") + 
  xlab("Age (year)")
pstaph2<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = St2)) +
  ylab("SB2 Relative Abundancy (%)") + 
  xlab("Age (year)")
pstaph1<-ggplot(data = data) + 
  geom_smooth(mapping = aes(x = age, y = St1)) +
  ylab("SB1 Relative Abundancy (%)") + 
  xlab("Age (year)")
```

 
```{r include = FALSE, message=FALSE}
###agregated species smoth line plots
SBR<-ggplot(data = data, aes(x = age, y = Staphylococcus)) +
  geom_point() +
  geom_smooth() +
  ylab("SB Relative Abundancy (%)") + 
  xlab("Age (year)")

CBR<-ggplot(data = data, aes(x = age, y = Corynebacterium)) + 
  geom_point() +
  geom_smooth() +
  ylab("CB Relative Abundancy (%)") + 
  xlab("Age (year)")

```

```{r, echo=FALSE, fig.height = 3, fig.width = 4, fig.align = "center", message=FALSE}
grid.arrange(SBR, CBR, nrow = 1)
```

For ease of interpretation and  comparison between  two Genus groups, we aggregate each of the four species in to one respective  group,  Corynebacterium and Staphylococcus. The result shows that the distribution of aggregated genus groups follows the dominant species distribution. As a result for Corynebacterium the relative abundancy gets minimum at the early age of around 20, and suddenly increases at around 24 year. It gets maximum approximately at age 43 and 60 years.  It drops gradually for ages older than 60 year. 

In case of Staphylococcus the relative abundancy is high at early age of about 20 and gradually deacreses with increasing age.  Lower relative abundacy for this genus is attined when age is between 40 and 60.  For older ages than 60 the relative abundancy starts to increase again.

It is clearly seen that the change in relative abundancy with age is not linear. Consequently, the correlations between the aggregated relative abundancy and age is calculated using the spearman correlation test in R.  correlation coefficient obtained for Corynebacterium and Staphylococcus is 0.40 and -0.41. since the association is not linear the correlation coefficient is small, but we got the information that Corynebacterium is positively associated while Staphylococcus is negatively associated. 

```{r include=FALSE}
#correlation test
cor.test(data$CB, data$age, method = "spearman")
cor.test(data$ST, data$age, method = "spearman")

```

## Relation of relative abundances of the 8 species to one another
To show that this test performs equivalent to a permutation test with n = 10000 (appendix: Hist and QQplot for a permutation result (n=10000) between CB1 & ST1), we performed a permutation test for the correlations between [Corynebacterium 1, Staphylococcus 1] and [Staphylococcus 3, Staphylococcus 4], these were chosen as in the former their distributions are not alike, while in the latter case they are more similar. From the results we can see that both permutation distributions are normally distributed, aside from the permutation distribution between Staph 3 and 4 having a small skew (appendix: Hist and QQplot for a permutation result (n=10000) between CB1 & ST1). The p-values are in the same range for both tests, for Cor 1 vs Staph 1, the t-test returned a p-value of 0.000334 and 0.00048 for the permutation. In the second comparison the results are 0.06973 and 0.0701 respectively. The asymptotic t approximation was used as this is a quicker method and the package allows for easy creation of confidence intervals and adjusting for multiple testing. 

The heatmaps for show that overall the correlation coefficients are small, the six significant pairs in the non-adjusted heatmap are [Cor 1 with Cor 2, Cor 3], [Cor 2 with Cor 3], [Staph 1 with Cor 1, Cor 2, Cor 3]. Adjusting for multiple testing leaves three significant pairs [Cor 1 with Cor 2] and [Staph 1 with Cor 1, Cor 2] (Figure 4). The confidence intervals however are very wide, therefore we cannot conclude anything about the magnitude of the association, only about the direction (Figure 5). From the significant pairs, Staph  1 is negatively correlated with Cor 1 and 2, while Cor 1 and 2 are positively associated with each other. A possible cause for this is if Staph 1 grows in the same microbial environment as Cor 1 and 2 and are therefore competitors. Cor 1 and 2 are positively associated, this could be because of either a symbiotic relation or because the environment in which they thrive is slightly different from one another, which allows them to grow together and not compete. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 5, fig.width = 6}
library(reshape2)
library(ggplot2)
library(ggpubr)

armpit <- data
#first 8 columns show the relative abundances, age, BMI and gender
# Get upper triangle
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
# Get lower triangle
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

library(psych)
#Create figure
a <- corr.test(armpit[1:8], method = "spearman", adjust = "BY")
#ccorrections possible: holm, hochberg, hommel, bonferroni, bh, by, fdr, none
#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045855/
cor_pvalues <- round(a$p,2)
cormat <- round(a$r,2)

#without adjusted p-values
lower_tri_p_values <- t(as.matrix(get_lower_tri(cor_pvalues)))
lower_tri <- t(as.matrix(get_lower_tri(cormat)))
# Melt the correlation matrix
melted_cormat <- melt(lower_tri, na.rm = TRUE)
melted_cormat_p_values_not_adjusted <- melt(lower_tri_p_values, na.rm = TRUE)

#lower_tri_p_values
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
  coord_fixed()

not_adjusted <- ggheatmap + 
  geom_text( aes(melted_cormat_p_values_not_adjusted$Var2, melted_cormat_p_values_not_adjusted$Var1, label = melted_cormat_p_values_not_adjusted$value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))


#With adjusted p-values
upper_tri_p_values <- as.matrix(get_upper_tri(cor_pvalues))
upper_tri <- as.matrix(get_upper_tri(cormat))
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
melted_cormat_p_values <- melt(upper_tri_p_values, na.rm = TRUE)

# Create a ggheatmap
ggheatmap2 <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
  coord_fixed()

adjusted <- ggheatmap2 + 
  geom_text( aes(melted_cormat_p_values$Var2, melted_cormat_p_values$Var1, label = melted_cormat_p_values$value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))

figure <- ggarrange(not_adjusted, adjusted,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)
#show(figure)
annotate_figure(figure, top = "Correlation test results \n A: p-values not adjusted for multiple testing \n B: p-values adjusted with the Benjamini-Yekutieli method")

```


# Conclusion
From the aggregated genus composition, the gene composition of the two groups change with age. in case of Staphylococcus genus, the relative abundancy is higher at early age which gradually decreases between 40 and 60 years of age, and it starts to increase after age 60. The opposite works for the Corynebacterium genus. However, in this analysis most of the people are younger than 40 year, and we recommend repeating the analysis in a homogenous data sample. 

## Genus composition among age categorical (agecat)
Since the p-value is greater than 0.05, we cannot to reject the null hypothesis, which means the genus composition do not shows significant difference among Agecat. 

## Genus composition for continuous age


## Relation of 8 relative abundances amongst each other


\clearpage


# References
Hauke, J., & Kossowski, T. (2011). Comparison of values of Pearson's and Spearman's correlation coefficients on the same sets of data. Quaestiones geographicae, 30(2), 87-93.

# Appendix

## Boxplots for the species relative abundance
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(reshape2)
library(ggplot2)
library(ggpubr)

armpit <- read.csv(paste0(getwd(),"/data/armpit.txt"), sep = ' ')

#first 8 columns show the relative abundances, age, BMI and gender
armpit <- armpit[complete.cases(armpit),]

armpit$Gender <- factor(trimws(armpit[["Gender"]], which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))
armpit$BMI <- factor(armpit$BMI)
#check if all relative abundances sum op to 100
#rowSums(armpit[1:8]) >= 99.99

#Check visually with boxplots and numerically with the kolmogorov-smirnov test 
#to see wether the distributions from two species are the same.

armpit_bxp1 <- melt(armpit[1:8])
bxp1 <- ggplot(armpit_bxp1, aes(x=variable, y=value)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1),
        axis.title.x = element_blank())+
  ylab("Relative abundances (%)") + 
  geom_boxplot()

armpit_bxp2 <- melt(log(armpit[1:8]+1))
bxp2 <- ggplot(armpit_bxp2, aes(x=variable, y=value)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1),
        axis.title.x = element_blank())+
  ylab("Log-transformed relative abundances") + 
  geom_boxplot()

figure <- ggarrange(bxp1, bxp2,
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

annotate_figure(figure, top = "Relative abundance boxplots for the 8 different species")

```

## Plots for each species versus age as continuous variable
```{r echo = FALSE, message=FALSE, warning=FALSE}
grid.arrange(pcor1, pcor2, pcor3, pcor4, nrow = 2)
grid.arrange(pstaph1, pstaph2, pstaph3, pstaph4, nrow = 2)
```
## Plots of difference of Spearman Rank correlation estimate significanes for different sample sizes
```{r spearmanbehav, echo=FALSE, message=FALSE, fig.height = 4, fig.width = 6}
p <- est <- px <- estx <- c()
for (i in 1:10000){
  Y1 <- rnorm(13)
  Y2 <- rnorm(13)
  Y1x <- rnorm(60)
  Y2x <- rnorm(60)
  p[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  est[i] <- cor.test(Y1,Y2, method = "spearman")$estimate
  px[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
  estx[i] <- cor.test(Y1x,Y2x, method = "spearman")$estimate
}
x <- data.frame(est, p)
xx <- data.frame(estx, px)
par(mfrow=c(1,2))
hist(x[x$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(x[x$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=13)'))
hist(xx[xx$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(xx[xx$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=60)'))
par(mfrow=c(1,1))
```

## Type1 simulation for Spearman rank
```{r spearmanType1, echo=FALSE, fig.height = 4, fig.width = 6, message=FALSE}
#H0 & Power testing for our groups
#assuming an even n in groups of 13 (39/3)
#sampling from runif and *100 to reflect out data
#Working with the aggregated level of CB & ST
##################
library(coin)
results_H0_A_runif <- data.frame(size=c(rep(26,6)),dist=rep('runif',6),
                               test=rep(c('perm_t','WMW','Spearman R'),2),
                               Test=c(rep('Type1',3),rep('Power',3)),
                               Value= 0)
pt_p <- wmw_p <- spr_p <- c()
p.t <- p.wmw <- p.spr <- c()
pt_pA <- wmw_pA <- spr_pA <- c()
p.tA <- p.wmwA <- p.sprA <- c()
n <- 13
for (i in 1:1000){
  #For H0= mean from same distrib
  Y1 <- runif(n)*100
  Y2 <- runif(n)*100
  #Groups
  X <- factor (c( rep("CB",n), rep("ST",n )))
  Y <- c(Y1 , Y2)
  #Power, now we need an actual difference
  Y1x <- runif(n, min = .40, max = .55)*100
  Y2x <- 100-Y1x
  #Groups
  Xx <- factor (c( rep("CB",n), rep("ST",n )))
  Yx <- c(Y1x , Y2x)
  # permutation H0
  p.t[i] <- pvalue(oneway_test(Y ~ X,
                                distribution = approximate (nresample =9999)))
  p.wmw[i] <- wilcox.test(Y1 ,Y2 ,
                          exact = TRUE )$p.value
  p.spr[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  # permutation HA
  p.tA[i] <- pvalue(oneway_test(Yx ~ Xx,
                               distribution = approximate (nresample =9999)))
  p.wmwA[i] <- wilcox.test(Y1x ,Y2x ,
                          exact = TRUE )$p.value
  p.sprA[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
}
# Monte - Carlo approximation of the power
# for alpha = .05
results_H0_A_runif$Value[1] <- mean(p.t < .05)
results_H0_A_runif$Value[2] <- mean(p.wmw < .05)
results_H0_A_runif$Value[3] <- mean(p.spr < .05)  
results_H0_A_runif$Value[4] <- mean(p.tA < .05)
results_H0_A_runif$Value[5] <- mean(p.wmwA < .05)
results_H0_A_runif$Value[6] <- mean(p.sprA < .05) 
library(ggplot2)
ggplot(results_H0_A_runif, aes(factor(dist), Value, fill = test)) + 
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Greys")+
  geom_hline(yintercept=0.05, col = 'red')+
  ggtitle("Power, Type1, for 10k reps")+
  facet_grid(. ~ Test)+
  theme_classic()
```

## Interaction of Agecat and BMI/Gender
```{r echo=FALSE, message=FALSE, fig.height = 4, fig.width = 6, fig.align = "center"}
#Interaction graphs
attach(data)
par(mfrow=c(1,2))
interaction.plot(Agecat, Gender, ST, 
                 xlab = "Agecat", 
                 ylab = "Relative Abundance of Staphylococcus", 
                 trace.label = "Gender",
                 main = "Interaction of Agecat and Gender",
                 ylim = c(0,100),  
                 pch = c(19,17), 
                 fixed = TRUE)
interaction.plot(Agecat, BMI, ST, 
                 xlab = "Agecat", 
                 ylab = "Relative Abundance of Staphylococcus", 
                 trace.label = "BMI",
                 main = "Interaction of Agecat and BMI",
                 ylim = c(0,100),  
                 pch = c(19,17), 
                 fixed = TRUE)
#Two-way ANOVA

```

## Boxplot of categorical age
```{r echo=FALSE, message=FALSE, fig.height = 3, fig.width = 4, fig.align = "center"}
#Kruskal Wallis test
attach(data)
boxplot(ST~Agecat,
        ylab = "Relative Abundance of Staphylococcus",
        main = "Staphylococcus by Agecat")

```

## Hist and QQplot for a permutation result (n=10000) between CB1 & ST1

```{r, include=FALSE, message=FALSE}
###Project Principle statistics###

library(reshape2)
library(ggplot2)
library(ggpubr)

armpit <- read.csv(paste0(getwd(),"/data/armpit.txt"), sep = ' ')
head(armpit)
summary(armpit)
#first 8 columns show the relative abundances, age, BMI and gender
armpit <- armpit[complete.cases(armpit),]

armpit$Gender <- factor(trimws(armpit[["Gender"]], which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))
armpit$BMI <- factor(armpit$BMI)
#check if all relative abundances sum op to 100
rowSums(armpit[1:8]) >= 99.99



ks_pvalues <- data.frame(matrix(nrow = dim(armpit)[2]-3, ncol = dim(armpit)[2] - 3))
rownames(ks_pvalues) <- colnames(armpit[1:8])
colnames(ks_pvalues) <- colnames(armpit[1:8])
for (x_index in 1:8){
  for (y_index in 1:8){
    x <- armpit[[x_index]]
    y <- armpit[[y_index]]
    ks_pvalues[x_index,y_index] <- ks.test(x,y)$p.value
  }
}

round(ks_pvalues,3)

#not all of the distributions are the same, therefore we permutate the order of the variables within each species

#Check correlation distribution

cor_permutation = c()
for (i in 1:10000){
  x = sample(armpit$Corynebacterium.1)
  y = sample(armpit$Staphylococcus.1)
  cor_permutation[i] = cor(x,y, method = "spearman")
}

test <- cor.test(armpit$Corynebacterium.1,armpit$Staphylococcus.1, method = "spearman" )
test$p.value
1 - mean(abs(cor_permutation) < abs(test$estimate))


# Get upper triangle
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
# Get lower triangle
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

library(psych)
#Create figure
a <- corr.test(armpit[1:8], method = "spearman", adjust = "BY")
#ccorrections possible: holm, hochberg, hommel, bonferroni, bh, by, fdr, none
#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045855/
cor_pvalues <- round(a$p,2)
cormat <- round(a$r,2)

#without adjusted p-values
lower_tri_p_values <- t(as.matrix(get_lower_tri(cor_pvalues)))
lower_tri <- t(as.matrix(get_lower_tri(cormat)))
# Melt the correlation matrix
melted_cormat <- melt(lower_tri, na.rm = TRUE)
melted_cormat_p_values_not_adjusted <- melt(lower_tri_p_values, na.rm = TRUE)

lower_tri_p_values
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
  coord_fixed()

not_adjusted <- ggheatmap + 
  geom_text( aes(melted_cormat_p_values_not_adjusted$Var2, melted_cormat_p_values_not_adjusted$Var1, label = melted_cormat_p_values_not_adjusted$value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))


#With adjusted p-values
upper_tri_p_values <- as.matrix(get_upper_tri(cor_pvalues))
upper_tri <- as.matrix(get_upper_tri(cormat))
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
melted_cormat_p_values <- melt(upper_tri_p_values, na.rm = TRUE)

# Create a ggheatmap
ggheatmap2 <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
  coord_fixed()

adjusted <- ggheatmap2 + 
  geom_text( aes(melted_cormat_p_values$Var2, melted_cormat_p_values$Var1, label = melted_cormat_p_values$value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))

figure <- ggarrange(not_adjusted, adjusted,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)
figure
annotate_figure(figure, top = "Correlation test results \n A: p-values not adjusted for multiple testing \n B: p-values adjusted with the Benjamini-Yekutieli method")

color_CI <- data.frame(t(a$ci.adj))
color_CI <- (colSums(color_CI > 0) == 2) | (colSums(color_CI < 0) == 2)
color_CI[color_CI == TRUE] <- "Red"
color_CI[color_CI == FALSE] <- "White"

rownames(a$ci.adj) <- rownames(a$ci)
par(mar=c(2,8,2,2))
boxplot(t(a$ci.adj), col = color_CI, horizontal = TRUE, main = "Adjusted confidence intervals", las = 1)
abline(v = 0)


```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
qq <- ggplot(data.frame(c = cor_permutation), aes(sample = c)) +
  stat_qq() + 
  stat_qq_line() +
  theme_minimal() +
  ylab("Frequency") + 
  xlab("Spearman Correlation")

histo <- ggplot(data.frame(c = cor_permutation), aes(x = c)) +
  geom_histogram(color="black", fill="white") + 
  theme_minimal() +
  ylab("Spearman Correlation") + 
  xlab("Theoretical Quantiles Normal Distribution")

figure <- ggarrange(histo, qq,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)
figure
annotate_figure(figure, top = "Histogram (A) and QQ-plot (B) for a permutation result (n = 10 000) \n between Corynebacterium 1 and Staphylococcus 1")


cor_permutation = c()
for (i in 1:10000){
  x = sample(armpit$Staphylococcus.3)
  y = sample(armpit$Staphylococcus.4)
  cor_permutation[i] = cor(x,y, method = "spearman")
}

test <- cor.test(armpit$Staphylococcus.3,armpit$Staphylococcus.4, method = "spearman" )
test$p.value
1 - mean(abs(cor_permutation) < abs(test$estimate))

qq <- ggplot(data.frame(c = cor_permutation), aes(sample = c)) +
  stat_qq() + 
  stat_qq_line() +
  theme_minimal() +
  ylab("Frequency") + 
  xlab("Spearman Correlation")

histo <- ggplot(data.frame(c = cor_permutation), aes(x = c)) +
  geom_histogram(color="black", fill="white") + 
  theme_minimal() +
  ylab("Spearman Correlation") + 
  xlab("Theoretical Quantiles Normal Distribution")

figure <- ggarrange(histo, qq,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1)
figure
annotate_figure(figure, top = "Histogram (A) and QQ-plot (B) for a permutation result (n = 10 000) \n between Staphylococcus 3 and Staphylococcus 4")
```

