---
title: "Principles of Statistical Analysis Report"
author: 
- Robin Boudry, Lin .., Merton..., Stijn...
date: "2019-12-26"
output:
  pdf_document:
    fig_caption: yes
    highlight: pygments
    keep_tex: yes
    number_sections: yes
    toc: no
header-includes: 
  - \usepackage{float}
  - \usepackage{longtable}
---
\centering

\raggedright

\clearpage

\tableofcontents


```{r, include=FALSE}
#INFO
#5 pages MAX
# no intro needed, just METHODS, RESULTS, CONCLUSION
# mention if assumptions are met for each test

```


# Methods

## Data Cleaning & Transformation
Data of 40 subjects was provided, containing information of relative abundances of 8 bacterial species (to be divided in 2 geni) present in their armpit, their BMI, age and gender. 1 observation was removed due to missing data. The categorical variable 'Gender' needed some cleaning to remove whitespaces around some codings of gendertype. For readability and easier coding the names 'Corynebacterium1' etc.. was shortened to 'Cb1' as was 'Staphylococcus1' etc.. shortened to 'St1'
A string variable 'top3' was created, refering to the top 3 classes per individual.
This gives us a 26 categories for 39 individuals...We also create top1 and top2, further aggregating the dominant bacteria and lessening the possible combinations.

### Descriptive analysis
When we view the data of age and CBorST, and sort on CBorST, we see no one above 30 has CB as dominant group. First indication of change in composition. When plotting the Geni with age, we see Corynebacterium having a cluster of low values in the younger age group, and Staphylococcus having a cluster of high values in the younger age group. 

```{r setup, echo=FALSE, message=FALSE, include=FALSE}
#'Setting up data & small cleaning & transformation'

#Pre Explorative: glancing
##################
data <- read.csv(paste0(getwd(),"/data/armpit.txt"), sep = ' ')
data <- data[complete.cases(data),]
data$Gender <- factor(trimws(data[["Gender"]], which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))
##################

#Rename the vars for readability
##################
names(data) <- c("Cb1", "Cb2", "Cb3", "Cb4", 
                 "St1", "St2", "St3", "St4",  
                 "Age", "BMI", "Gender" )
##################

#Create top 3 classes per observations
##################
#x <- toString(names(sort(data[1,1:8],decreasing = TRUE)[1:3]))
#top3
data$top3 <- 0
for(i in 1:nrow(data)){
  data$top3[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:3]))
}
#table(data$top3)

#We see 26 combinations for 39 observations.... With this data any difference
#test will seem significant.
#Let's build top 2

#top2
data$top2 <- 0
for(i in 1:nrow(data)){
  data$top2[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1:2]))
}
#table(data$top2)
#We still see 13 categories here... Still a lot.
#Lets build top 1

#top1
data$top1 <- 0
for(i in 1:nrow(data)){
  data$top1[i] <- toString(names(sort(data[i,1:8],decreasing = TRUE)[1]))
}
#table(data$top1)
#Here we have only 4 categories. However the question was composition.
#Let's aggregate Cv and ST, and then build top 2 & top 3

#aggregation
data$CB <- 0
data$ST <- 0
for(i in 1:nrow(data)){
  data$CB[i] <- sum(data[i,1:4])
  data$ST[i] <- sum(data[i,5:8])
}
par(mfrow=c(1,2))
plot(data$Age, data$CB, main = 'CB & Age')
plot(data$Age, data$ST, main = 'ST & Age')
par(mfrow=c(1,1))
#We get a hint of increase in CB, and decrease in ST, but needs testing.

#Let's build CB_ST:
#Which is dominant?
data$CBorST <- 0
for(i in 1:nrow(data)){
  data$CBorST[i] <- toString(names(sort(data[i,15:16],decreasing = TRUE)[1:2]))
}
#table(data$CBorST)

#When we view the dataset, and sort on CBorST, we see no one above 30 has CB
#as dominant group. First indication of change in composition
#shown with dplyr
library(dplyr)
#data%>%
#  select(Age, CBorST)%>%
#  filter(CBorST == "CB, ST")

##################


```


## Choice of non parametric tests
We are confronted with asample size and since no information is given about the underlying distribution, we have to rely on non-parametric methods to make valid inferences.
For correlation testing, we consulted Hauke & Kowoski(2011)'s comparison of correlation tests, and concluded that Spearman rank was suitable for our problem.

To better understand the Spearman Rank correlation significance test, we simulated
data with the same sample size (n1=n2=13) which resembles what we observed and with higher sample sizes.. We noted that smaller sample sizes take higher correlation estimates for a significant p-value (see appendix).
We also checked the Type1 error in a simulation, and saw it being controlled for Spearman Rank cor test, similar to more established tests (see appendix). 

For group comparisons the anova was out of order, but the Wilcoxon Rank test as seen in the course was suitable.


# Results

# Conclusion

# References
Hauke, J., & Kossowski, T. (2011). Comparison of values of Pearson's and Spearman's correlation coefficients on the same sets of data. Quaestiones geographicae, 30(2), 87-93.

# Appendix
A look at difference of Spearman Rank correlation estimate significanes for different sample sizes
```{r spearmanbehav, echo=FALSE, message=FALSE}
p <- est <- px <- estx <- c()
for (i in 1:10000){
  Y1 <- rnorm(13)
  Y2 <- rnorm(13)
  Y1x <- rnorm(60)
  Y2x <- rnorm(60)
  p[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  est[i] <- cor.test(Y1,Y2, method = "spearman")$estimate
  px[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
  estx[i] <- cor.test(Y1x,Y2x, method = "spearman")$estimate
}
x <- data.frame(est, p)
xx <- data.frame(estx, px)
par(mfrow=c(1,2))

hist(x[x$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(x[x$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=13)'))

hist(xx[xx$p<=0.05,1], breaks = 40, main=paste('+-', round(min(abs(xx[xx$p<0.05,1])),3),
     'seems to be minimum Spearman Rank \n cor value for 0.05 p-value \n (n1=n2=60)'))

par(mfrow=c(1,1))
```

Type1 simulation for Spearman rank.
```{r spearmanType1, echo=FALSE, message=FALSE}

#H0 & Power testing for our groups
#assuming an even n in groups of 13 (39/3)
#sampling from runif and *100 to reflect out data
#Working with the aggregated level of CB & ST
##################
library(coin)
results_H0_A_runif <- data.frame(size=c(rep(26,6)),dist=rep('runif',6),
                               test=rep(c('perm_t','WMW','Spearman R'),2),
                               Test=c(rep('Type1',3),rep('Power',3)),
                               Value= 0)
pt_p <- wmw_p <- spr_p <- c()
p.t <- p.wmw <- p.spr <- c()
pt_pA <- wmw_pA <- spr_pA <- c()
p.tA <- p.wmwA <- p.sprA <- c()
n <- 13
for (i in 1:1000){
  #For H0= mean from same distrib
  Y1 <- runif(n)*100
  Y2 <- runif(n)*100
  #Groups
  X <- factor (c( rep("CB",n), rep("ST",n )))
  Y <- c(Y1 , Y2)
  #Power, now we need an actual difference
  Y1x <- runif(n, min = .40, max = .55)*100
  Y2x <- 100-Y1x
  #Groups
  Xx <- factor (c( rep("CB",n), rep("ST",n )))
  Yx <- c(Y1x , Y2x)
  # permutation H0
  p.t[i] <- pvalue(oneway_test(Y ~ X,
                                distribution = approximate (nresample =9999)))
  p.wmw[i] <- wilcox.test(Y1 ,Y2 ,
                          exact = TRUE )$p.value
  p.spr[i] <- cor.test(Y1,Y2, method = "spearman")$p.value
  # permutation HA
  p.tA[i] <- pvalue(oneway_test(Yx ~ Xx,
                               distribution = approximate (nresample =9999)))
  p.wmwA[i] <- wilcox.test(Y1x ,Y2x ,
                          exact = TRUE )$p.value
  p.sprA[i] <- cor.test(Y1x,Y2x, method = "spearman")$p.value
}
# Monte - Carlo approximation of the power
# for alpha = .05


results_H0_A_runif$Value[1] <- mean(p.t < .05)
results_H0_A_runif$Value[2] <- mean(p.wmw < .05)
results_H0_A_runif$Value[3] <- mean(p.spr < .05)  
results_H0_A_runif$Value[4] <- mean(p.tA < .05)
results_H0_A_runif$Value[5] <- mean(p.wmwA < .05)
results_H0_A_runif$Value[6] <- mean(p.sprA < .05) 

library(ggplot2)
ggplot(results_H0_A_runif, aes(factor(dist), Value, fill = test)) + 
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Greys")+
  geom_hline(yintercept=0.05, col = 'red')+
  ggtitle("Power, Type1, for 10k reps")+
  facet_grid(. ~ Test)+
  theme_classic()
```

